---
title: "QFishData_OtterTrawls"
author: "Hannah McClive"
date: "2025-04-09"
output: html_document
---

Assignment. End-to-end data analysis in R.

In this assignment, we will use real-world marine data collected as part of Queensland fisheries QFISH database to develop
a report (either written on your website or via an R Markdown report embedded in your website) suitable for reporting to 
your manager, a community group or for your own research.

Part 1. First, obtain your data. 
Go to QFISH and review all of the available datasets that you could analyse. 
Here you will note that there are summaries available - I suggest a dataset that has 
both summaries and catch information in space (like regions) and by species. 
For example, selecting the Shark Control Program tab at left, then selecting the number 
caught by area, calendar year and species group offers a data table by species, year and region.

Identify the dataset you want to analyse
Export the dataset in an appropriate format for R.
Use your Github repo to save the data in an appropriate place.
Ensure you sync your github regularly by committing and pushing to the remote repository.

Part 2. Tidy the data.
Using the tidyverse, get your data ready for making a plot.
Use R Markdown to write your analysis up as a report that shows all of the steps 
you used to import the data and plot it. 
(Note: it’s ok if you want to simply use a script - in this case build your report directly on your website).

Develop a time-series (or any other plot) that you want to make.
For some useful advice, you can use the resources provided earlier in this book, 
or the excellent ggplot2 page of the Ocean Health Index website.
Export your plot in a format suitable to put on your e-portfolio.

Part 3. Add your plot and markdown report to your e-portfolio
In a section of your website, add your full report that details your analysis, 
where you got the data, and the analysis you developed (in sections, follow good coding style!) 
to get there.
Finally, include your plot together with a detailed caption of what your plot shows.
Your final e-portfolio should be suitable for sharing with a colleague or the public, 
so that they can understand the dataset you have presented to them.

```{r}
# Loading and renaming QFish data (imported as excel file)
QFishData <- MB5370_Mod4_QFish
```

```{r}
# Installing and checking for relevant packages
library(tidyverse)
```

```{r}
# Clean column names 
QFishData <- QFishData %>%
  rename(
    Year = CalendarYear,
    Catch_tonnes = Tonnes,
    Fishing_days = Days
  ) %>%
  filter(!is.na(Year)) %>%
  mutate(Year = as.numeric(Year))
```
```{r}
library(tidyverse)
library(readxl)

#QFishData <- read_excel("MB5370_Mod4_QFish.xlsx") %>%
 # rename(
  #  Year = CalendarYear,
   # Catch_tonnes = Tonnes,
    #Fishing_days = Days
 # )

# Check for non-numeric values in Year
print(unique(QFishData$Year))

# Force Year to numeric and remove problem rows
QFishData <- QFishData %>%
  mutate(Year = as.numeric(Year)) %>%
  filter(!is.na(Year))

# Confirm it worked
print(range(QFishData$Year))
```

```{r}
QFishFiltered <- QFishData %>%
  filter(Catch_tonnes < 100000)
```

```{r}
ggplot(QFishFiltered, aes(x = Year, y = Catch_tonnes)) +
  geom_line(color = "steelblue", size = 1.2) +
  geom_point(color = "darkblue") +
  scale_x_continuous(
    breaks = seq(min(QFishFiltered$Year), max(QFishFiltered$Year), by = 2)
  ) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    title = "Total Catch Over Time",
    x = "Year",
    y = "Catch (Tonnes)",
    caption = "Source: QFish Dataset"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", size = 16)
  )
```

